Forecasting interrupted time series

### 2023-07-28 fits.qmd
More work on the tourism example

# Load main data objects from targets folder
tar_config_set(store = here::here("_targets"))
```

@@

The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia. The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1. As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021. The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that. The data are available from the Australian Bureau of Statistics.

@@

We will apply each of the approaches above to this data set, and compare their performance.

```{r}
#| label: fig-tsol1_plot
#| fig-height: 8
#| fig-cap: Solution 1. This should show that ETS is highly adaptive. However, for some reason, that is no longer the case, and I don't know why.
tar_read(tsol1_plot)
```

```{r}
#| label: fig-tsol2_plot
#| fig-height: 8
#| fig-cap: Solution 2. We use a left shift in March 2020, and a ramp from November 2022 to June 2023.
tar_read(tsol2_plot)
```

```{r}
#| label: fig-tsol3_plot
#| fig-height: 8
#| fig-cap: Solution 3. The period from March 2020 to November 2022 is set to missing. So the first three years of forecasts show what might have been. The fourth set of forecasts uses observations from November and December 2022, and so the forecasts have been adjusted downwards.
tar_read(tsol3_plot)
```

```{r}
#| label: fig-tsol4_plot
#| fig-height: 8
#| fig-cap: Solution 4. We replace the observations from March 2020 to November 2022 with the average of the same month in the 3 preceding years. 
tar_read(tsol4_plot)
```

>>>>

# Load main data objects from targets folder
tar_config_set(store = here::here("_targets"))
tar_load(c(austa))
```

@@

The Australian tourism data set is a monthly time series showing the number of short-term overseas visitors to Australia. The data [@tourismdata] are available from January 2000 to May 2023, and are shown in @fig-tourism_plot1. As the borders closed in March 2020, the number of visitors to Australia dropped to near zero, and remained there until towards the end of 2021. The borders officially reopened on 21 February 2022, although it seems visitors began to arrive earlier than that.

@@

We apply the four solutions discussed in the previous section to these data, making 12 month forecasts at the end of each year from 2019 to 2022. We fit ETS, ARIMA and dynamic ARIMA models to the data [@fpp3], first applying a log transformation to ensure the resulting forecasts are positive. This is possible, because the observations never reach exactly zero, with the smallest number of visitors per month equal to `r min(austa$Visitors)*1000` in `r austa |> filter(Visitors == min(Visitors)) |> pull(Month) |> format("%B %Y")`. For all forecasts, we also show 90% prediction intervals.

```{r}
#| label: fig-tsol1_plot
#| fig-height: 8
#| fig-pos: "t"
#| fig-cap: Solution 1. This should show that ETS is highly adaptive. However, for some reason, that is no longer the case, and I don't know why.
tar_read(tsol1_plot)
```

In @fig-tsol1_plot, we show forecasts by applying ETS and ARIMA models to the data. ETS, in particular, is well-known to be relatively adaptive to changes in the series, and this is evident in these forecasts. The forecasts for 2020 were made using data before COVID-19 had any affect, and so they show similar patterns to the past. The forecasts for 2021 were made after 9 months of very low levels of arrivals, and both models show forecasts consistent with the recent history. The use of logarithms is particularly important here as the variance is much smaller during 2020 than previously, but after taking logarithms, the variance is more stable over time. The forecasts made at the start of 2022 have struggled to detect the small increase in traffic at the end of 2021, and both models have forecast relatively flat trajectories as a result, although the prediction intervals are wide indicating model uncertainty. Finally, the forecasts made at the end of 2022 have captured the increasing trend, but ETS is much closer to the reality, adapting more quickly to the changing patterns. Again, the wide prediction intervals indicate a high level of uncertainty. It is possible to make ETS more adaptive to changes in the data by increasing the value of the smoothing parameters. For example, a high value of $\beta$ (the smoothing parameter for the slope), will result in changes of trend being incorporated into the forecasts more quickly, at the risk of over-reacting to noise in the data.

```{r}
#| label: fig-tsol2_plot
#| fig-height: 8
#| fig-pos: "t"
#| fig-cap: Solution 2. We use a left shift from March 2020 to October 2022, and a ramp from October 2021 to October 2022. After October 2022, there are no non-zero predictors.
tar_read(tsol2_plot)
```

@fig-tsol2_plot shows forecasts obtained using a dynamic regression model (i.e., a regression with ARIMA errors) using two intervention variables: a level shift from March 2022 to November 2022, and a ramp from October 2021 to November 2022. It is evident that the level shift variable has worked well, giving relatively good forecasts for 2021. However, the forecasts for 2022 are particularly poor, because the ramp slope has been greatly overestimated, as it was based on only three observations (Oct -- Dec 2021). The forecasts for 2023 are much better, and the relatively large prediction intervals are appropriate given the uncertainty in the industry at the end of 2022.

```{r}
#| label: fig-tsol3_plot
#| fig-height: 8
#| fig-pos: "t"
#| fig-cap: Solution 3. The period from March 2020 to October 2022 is set to missing. So the first three years of forecasts show what might have been. The fourth set of forecasts uses observations from November and December 2022, and so the forecasts have been adjusted downwards.
tar_read(tsol3_plot)
```

The third solution involved setting the observations during the disruption period to missing, and then fitting an ARIMA model to the series. Consequently, the first three years of forecasts show what might have been without the COVID-19 pandemic, based on the history to the end of 2019. The final forecasts for 2023 use the data to the end of 2019, and the two observations in November and December 2022. These are much better, but the prediction intervals are too narrow, as there hasn't been sufficient recent data.

```{r}
#| label: fig-tsol4_plot
#| fig-height: 8
#| fig-pos: "t"
#| fig-cap: Solution 4. We replace the observations from March 2020 to October 2022 with the average of the same month in the three years prior to March 2020.
tar_read(tsol4_plot)
```

Finally, we show in @fig-tsol4_plot the forecasts obtained using solution 4. Here we have replaced observations between March 2020 and October 2022 with estimates based on the average of the same month in the three years prior to March 2020. The resulting forecasts are similar to those from solution 3, but with narrower prediction intervals because the model is (falsely) assuming that the "observations" during the disruption are real.

### 2023-07-28 fits.tex
More work on the tourism example

\usepackage[section]{placeins}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
\makeatother
\makeatletter
\makeatother

@@

The Australian tourism data set is a monthly time series showing the
number of short-term overseas visitors to Australia. The data
\autocite{tourismdata} are available from January 2000 to May 2023, and
are shown in Figure~\ref{fig-tourism_plot1}. As the borders closed in
March 2020, the number of visitors to Australia dropped to near zero,
and remained there until towards the end of 2021. The borders officially
reopened on 21 February 2022, although it seems visitors began to arrive
earlier than that. The data are available from the Australian Bureau of
Statistics.

@@

We will apply each of the approaches above to this data set, and compare
their performance.

@@

@@

\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift in March
2020, and a ramp from November 2022. We possibly need to stop the ramp
around mid 2023.}

@@

@@

\caption{\label{fig-tsol3_plot}Solution 3. The period from March 2020 to
November 2022 is set to missing. So the first three years of forecasts
show what might have been. The fourth set of forecasts uses observations
from November and December 2022, and so the forecasts have been adjusted
downwards.}

@@

@@

\caption{\label{fig-tsol4_plot}Solution 4. We replace the observations
from March 2020 to November 2022 with the average of the same month in
the 3 preceding years.}

@@

>>>>

\usepackage[section]{placeins}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
\makeatother
\makeatletter
\makeatother

@@

The Australian tourism data set is a monthly time series showing the
number of short-term overseas visitors to Australia. The data
\autocite{tourismdata} are available from January 2000 to May 2023, and
are shown in Figure~\ref{fig-tourism_plot1}. As the borders closed in
March 2020, the number of visitors to Australia dropped to near zero,
and remained there until towards the end of 2021. The borders officially
reopened on 21 February 2022, although it seems visitors began to arrive
earlier than that.

@@

We apply the four solutions discussed in the previous section to these
data, making 12 month forecasts at the end of each year from 2019 to
2022. We fit ETS, ARIMA and dynamic ARIMA models to the data
\autocite{fpp3}, first applying a log transformation to ensure the
resulting forecasts are positive. This is possible, because the
observations never reach exactly zero, with the smallest number of
visitors per month equal to 2250 in April 2020.

\begin{figure}[t]

@@

In Figure~\ref{fig-tsol1_plot}, we show forecasts by applying ETS and
ARIMA models to the data. ETS, in particular, is well-known to be
relatively adaptive to changes in the series, and this is evident in
these forecasts. The forecasts for 2020 were made using data before
COVID-19 had any affect, and so they show similar patterns to the past.
The forecasts for 2021 were made after 9 months of very low levels of
arrivals, and both models show forecasts consistent with the recent
history. The use of logarithms is particularly important here as the
variance is much smaller during 2020 than previously, but after taking
logarithms, the variance is more stable over time. The forecasts made at
the start of 2022 have struggled to detect the small increase in traffic
at the end of 2021, and both models have forecast relatively flat
trajectories as a result, although the prediction intervals are wide
indicating model uncertainty. Finally, the forecasts made at the end of
2022 have captured the increasing trend, but ETS is much closer to the
reality, adapting more quickly to the changing patterns. Again, the wide
prediction intervals indicate a high level of uncertainty. It is
possible to make ETS more adaptive to changes in the data by increasing
the value of the smoothing parameters. For example, a high value of
\(\beta\) (the smoothing parameter for the slope), will result in
changes of trend being incorporated into the forecasts more quickly, at
the risk of over-reacting to noise in the data.

\begin{figure}[t]

@@

\caption{\label{fig-tsol2_plot}Solution 2. We use a left shift from
March 2020 to October 2022, and a ramp from October 2021 to October
2022. After October 2022, there are no non-zero predictors.}

@@

Figure~\ref{fig-tsol2_plot} shows forecasts obtained using a dynamic
regression model (i.e., a regression with ARIMA errors) using two
intervention variables: a level shift from March 2022 to November 2022,
and a ramp from October 2021 to November 2022. It is evident that the
level shift variable has worked well, giving relatively good forecasts
for 2021. However, the forecasts for 2022 are particularly poor, because
the ramp slope has been greatly overestimated, as it was based on only
three observations (Oct -- Dec 2021). The forecasts for 2023 are much
better, and the relatively large prediction intervals are appropriate
given the uncertainty in the industry at the end of 2022.

\begin{figure}[t]

@@

\caption{\label{fig-tsol3_plot}Solution 3. The period from March 2020 to
October 2022 is set to missing. So the first three years of forecasts
show what might have been. The fourth set of forecasts uses observations
from November and December 2022, and so the forecasts have been adjusted
downwards.}

@@

The third solution involved setting the observations during the
disruption period to missing, and then fitting an ARIMA model to the
series. Consequently, the first three years of forecasts show what might
have been without the COVID-19 pandemic, based on the history to the end
of 2019. The final forecasts for 2023 use the data to the end of 2019,
and the two observations in November and December 2022. These are much
better, but the prediction intervals are too narrow, as there hasn't
been sufficient recent data.

\begin{figure}[t]

@@

\caption{\label{fig-tsol4_plot}Solution 4. We replace the observations
from March 2020 to October 2022 with the average of the same month in
the 3 preceding years.}

@@

Finally, we show in Figure~\ref{fig-tsol4_plot} the forecasts obtained
using solution 4. Here we have replaced observations between March 2020
and October 2022 with estimates based on the average of the same month
in the three preceding years. The resulting forecasts are similar to
those from solution 3, but with narrower prediction intervals because
the model is (falsely) assuming that the ``observations'' during the
disruption are real.
